Bayesian model in JAGS

### 2023-01-20 0_HCUP_data_clean/Data directionary.Rmd
change file names to show order

  - \textcolor{red}{HCUPCode}. It contains the sas codes to read in the asc files from HCUP data.
  \
  \
  - \textcolor{red}{HCUPData}. It contains the original asc data files from the zip files you decompress from CD.
  \
  \
  - \textcolor{red}{ParquetData}. This is the final data files you will use for analysis. Each state has two files. One file contains the core data of all years, including diagnoses, date of birth, date of admission etc. The other file named chgs contains the charges information (useful for cost-effectiveness studies). 
  \
  \
  - \textcolor{red}{SASCode}. These are the codes that Iris and I (Gigi) wrote to convert all the asc files to csv files.
  \
  \
  - \textcolor{red}{SASData}. This contains the output csv files from the SAS program.

@@

(1) Read in CD. You need to ensure your computer have at least 200G space for storing the data.
  \
  \
(2) Uncompressed the zip files by inputting the passwords for each year and each state.
  \
  \
(3) Copy the asc files to directory \textcolor{red}{HCUPData}.
  \
  \
(4) Run the *sid time series.sas* in \textcolor{red}{SASCode} to run the SAS codes in \textcolor{red}{HCUPCode}. This will load all the asc data into SAS environment. And then run the *ConvertCSV.sas* to convert all data in SAS environment into csv files.
  \
  \
(5) Download R file: https://github.com/Gigi112/MedicareCMSExploration/blob/main/2.CreateParquet_states.Rmd and modify it to convert all csv files to Parquet files.
  \
  \
(6) Modify R file: https://github.com/VA-CareDisruptions/MedicareCMSExploration/blob/main/1.%20exploratory%20analyses.Rmd to analyze parquet data in R environment.

>>>>

  - \textcolor{red}{HCUPCode}. It contains the sas codes to read in the asc files from HCUP data.
  \
  \
  - \textcolor{red}{HCUPData}. It contains the original asc data files from the zip files you decompress from CD.
  \
  \
  - \textcolor{red}{ParquetData}. This is the final data files you will use for analysis. Each state has two files. One file contains the core data of all years, including diagnoses, date of birth, date of admission etc. The other file named chgs contains the charges information (useful for cost-effectiveness studies). 
  \
  \
  - \textcolor{red}{SASCode}. These are the codes that Iris and I (Gigi) wrote to convert all the asc files to csv files.
  \
  \
  - \textcolor{red}{SASData}. This contains the output csv files from the SAS program.
    \
  \
  - \textcolor{red}{Rcode}. This contains the R codes to convert csv files to Parquet files. It also contains the R codes to run exploratory analyses using Parquet data.
    \
  \

@@

(1) Read in CD. You need to ensure your computer have at least 200G space for storing the data.
  \
  \
(2) Uncompressed the zip files by inputting the passwords for each year and each state.
  \
  \
(3) Copy the asc files to directory \textcolor{red}{HCUPData}.
  \
  \
(4) Run the *sid time series.sas* in \textcolor{red}{SASCode} to run the SAS codes in \textcolor{red}{HCUPCode}. This will load all the asc data into SAS environment. And then run the *ConvertCSV.sas* to convert all data in SAS environment into csv files.
  \
  \
(5) Within \textcolor{red}{Rcode}, run R code *0_2.CreateParquet_states.Rmd* and modify it to convert all csv files to Parquet files.
  \
  \
(6) Within \textcolor{red}{Rcode}, modify R code *1_2.exploratory analyses.Rmd* to analyze parquet data in R environment.

### 2023-01-20 3_JAGS_Bayesian_model/Jags_Bayesian_model.Rmd
change file names to show order

```{r, eval=FALSE}
################ priors ####################################
beta0~dnorm(0,0.0001)
beta1~dnorm(0,0.0001)
lambda1~dnorm(0,0.0001)
lambda2~dnorm(0,0.0001)
inv_var ~ dgamma(0.01, 0.01)}"
```
\
\
Since the model running process is the same, I will skip the model fitting process in this tutorial (it takes time to knit the PDF). To run the model, you need to combine the two chunks of model structures and change the eval from False to True.

@@

>>>>

```{r, eval=FALSE}
"model{# delete this one line of code
# when combine the two chunks. Separating the
# codes helps to show them in slides

################ priors ####################################
beta0~dnorm(0,0.0001)
beta1~dnorm(0,0.0001)
lambda1~dnorm(0,0.0001)
lambda2~dnorm(0,0.0001)
inv_var ~ dgamma(0.01, 0.01)}"
```
\
\
Since the model running process is the same, I will skip the model fitting process in this tutorial (it takes time to knit the PDF). To run the model, you need to combine the two chunks of model structures and change the eval from False to True.

@@

Very often, the disease incidence is not randomly distributed across space. Observations in nearby localities tend to be more similar than observations further away, also known as positive spatial autocorrelation. 
\
\
Prof. Duncan Lee has developed many statistical tools and tutorials to help evaluate spatial autocorrelation in public health. Please refer to the following materials to have an complete overview on Bayesian spatial model for public health and choosing conditional autoregressive priors: 
\
\

- [Bayesian Disease Mapping for Public Health](./Bayesian Disease Mapping for Public Health)

- [Overview of Spatial Incidence Models](https://atlas.cancer.org.au/developing-a-cancer-atlas/Chapter_5.html)

- [CARBayes: An R Package for Bayesian Spatial Modeling with Conditional Autoregressive Priors](./CARBayes: An R Package for Bayesian Spatial Modeling with Conditional Autoregressive Priors.pdf)

- [Spatio-Temporal Areal Unit Modeling in R with Conditional Autoregressive Priors Using the CARBayesST Package](./Spatio-Temporal Areal Unit Modeling in R with Conditional Autoregressive Priors Using the CARBayesST Package.pdf)

## Spatial model with censored data
```{r, eval=FALSE}
censor_spatial_model <- "model{
for(i in 1:n){
is.censor[i] ~ dinterval(Z[i],t.cen)
Z[i] ~ dlnorm(mu[i], inv_var)
 mu[i] <- beta0[i] + beta1*log_W[i]
}

beta0[1:n] ~ dmnorm(intercepts, Leroux)
 for(i in 1:n){
 intercepts[i] <- b0+x1[i]*b1+x2[i]*b2}

beta1~dnorm(0,0.0001)

b0~dnorm(0,0.0001)
b1~dnorm(0,0.0001)
b2~dnorm(0,0.0001)
```

## Spatial model with censored data (continued)
```{r, eval=FALSE}
"model{# delete this one line of code
# when combine the two chunks. Separating the
# codes helps to show them in slides

Leroux <- inv.tau*(rho*w.mat + (1 - rho)*Id.mat) 
# calculate spatial correlated random effects
  inv.tau ~ dgamma(0.01, 0.01) 
# note the w.mat here is equal to 
# diag(rowSums(neighbors_mat)) - neighbors_mat

  rho~ dunif(0,1)
inv_var ~ dgamma(0.01, 0.01)}"
```
